Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-06-10 covid_nrs_place_of_death_visualisations_2022.Rmd
added cause of death plots for data sicne 2020 and by year; also added a line that hopefulyl stops the annyoing messages about missing fonts

```{r}
loadfonts()  # load imported fonts
# Sys.setenv(R_GSCMD = "C:/Program Files/gs/gs9.27/bin/gswin64c.exe")  # set the location of the GS file  # this not needed, just use device=device_cairo for pdfs in ggsave

theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri")
            ))

@@

```{r}
## to put the arrows explaining historic range etc. only in one panel, need to construct a tibble to pass as data
annotations_figure_place_of_death_explaining_historic_range <- tribble(
   ~geom, ~place_of_death, ~label, ~x, ~y, ~xend, ~yend,
  "text", "Hospital", "Historic range (min & max)", 29, 850, NA, NA,
  "curve", "Hospital", NA, 29, 830, 32, 550,
  "text", "Hospital", "Historic mean", 22, 800, NA, NA,
  "curve", "Hospital", NA, 22, 780, 26, 510
) %>% 
  mutate(across(.cols = c("x","xend"), .fns = ~compute_start_date_from_week_number(week_number = .x, year_number = 2020))) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home")))  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)

figure_place_of_death <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(!place_of_death %in% c("All","Other")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  geom_text(data = annotations_figure_place_of_death_explaining_historic_range %>% filter(geom=="text"), aes(x=x, y=y, label=label),  colour = "grey20", size = 8 / (14/5)) +  # the size is tricky
  geom_curve(data = annotations_figure_place_of_death_explaining_historic_range %>% filter(geom=="curve"), aes(x=x, y=y, xend=xend, yend=yend), colour = c("#4477aa", "grey20"), alpha = c(0.5, 1), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 28, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
figure_place_of_death_home_only <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2022_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death_home_only %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_only", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```

@@

figure_place_of_death_home_2015_2019_historic_period <-
  merged_deaths_overall_2015_2019_historic_period %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2015_2019_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "2022", hjust = 0, colour = "grey40") +
  NULL

figure_place_of_death_home_2015_2019_historic_period %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_2015-2019_range", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
figure_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

figure_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

```{r}
figure_weekly_proportion_of_death_by_place <-
  bind_rows(
    ## current data
    merged_deaths_overall %>%
      filter(age == "all" & sex == "all"),
    ## 2015-2019
    weekly_deaths_2015_2019_overall %>%
      mutate(year = as.integer(year)) %>%  # remove the 'average' figures
      drop_na(year) %>%
      rename(deaths_all_causes = number_of_deaths)
  ) %>%
  group_by(year, week_number, place_of_death == "All") %>%
  mutate(proportion_of_deaths = deaths_all_causes / sum(deaths_all_causes)) %>%
  ungroup %>%
  filter(place_of_death != "All") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(
    data = ., aes(x = week_number, y = proportion_of_deaths, colour = place_of_death)
  ) +
  geom_line() +
  facet_wrap(~year) +
  scale_colour_viridis_d() +
  scale_x_continuous(breaks = scales::breaks_extended(6)) +
  scale_y_continuous(breaks = scales::breaks_extended(6), label = function(x) scales::percent(x = x, accuracy = 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
  ) +
  labs(
    x = "Week number", y = "Weekly proportion of deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
  ) +
  NULL

figure_weekly_proportion_of_death_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_weekly_proportion_of_death_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 20, height = 15, dpi = 300, units = "cm")
```

@@

```{r}
figure_cumulative_weekly_proportion_of_deaths_by_place <-
  cumulative_deaths_by_place_by_year %>%
  filter(place_of_death != "All") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(data = ., aes(x = week_number, y = proportion_cumulative_deaths, group = place_of_death, colour = place_of_death)) +
  geom_line() +
  facet_wrap(~year) +
  scale_colour_viridis_d() +
  scale_fill_viridis_d() +
  scale_x_continuous(breaks = scales::breaks_extended(6)) +
  scale_y_continuous(breaks = scales::breaks_extended(6), label = function(x) scales::percent(x = x, accuracy = 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
  ) +
  labs(
    x = "Week number", y = "Cumulative proportion of place of death (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
  ) +
  NULL

figure_cumulative_weekly_proportion_of_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_cumulative_weekly_proportion_of_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 20, height = 15, dpi = 300, units = "cm")
```

>>>>

```{r}
loadfonts(device="win")  # load imported fonts
# Sys.setenv(R_GSCMD = "C:/Program Files/gs/gs9.27/bin/gswin64c.exe")  # set the location of the GS file  # this not needed, just use device=device_cairo for pdfs in ggsave
windowsFonts("Calibri" = windowsFont("Calibri"))  # to suppress the multiple error messages

theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri"),
                  axis.text = element_text(size = 9)
            ))

@@

```{r}
## to put the arrows explaining historic range etc. only in one panel, need to construct a tibble to pass as data
annotations_fig_place_of_death_explaining_historic_range <- tribble(
   ~geom, ~place_of_death, ~label, ~x, ~y, ~xend, ~yend,
  "text", "Hospital", "Historic range (min & max)", 29, 850, NA, NA,
  "curve", "Hospital", NA, 29, 830, 32, 550,
  "text", "Hospital", "Historic mean", 22, 800, NA, NA,
  "curve", "Hospital", NA, 22, 780, 26, 510
) %>% 
  mutate(across(.cols = c("x","xend"), .fns = ~compute_start_date_from_week_number(week_number = .x, year_number = 2020))) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home")))  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)

fig_place_of_death <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(!place_of_death %in% c("All","Other")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  geom_text(data = annotations_fig_place_of_death_explaining_historic_range %>% filter(geom=="text"), aes(x=x, y=y, label=label),  colour = "grey20", size = 8 / (14/5)) +  # the size is tricky
  geom_curve(data = annotations_fig_place_of_death_explaining_historic_range %>% filter(geom=="curve"), aes(x=x, y=y, xend=xend, yend=yend), colour = c("#4477aa", "grey20"), alpha = c(0.5, 1), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

fig_place_of_death %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 28, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
fig_place_of_death_home_only <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2022_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "'22", hjust = 0, colour = "grey40") +
  NULL

fig_place_of_death_home_only %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_only", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```

@@

fig_place_of_death_home_2015_2019_historic_period <-
  merged_deaths_overall_2015_2019_historic_period %>%
  filter(sex == "all" & age == "all") %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%  # home death only
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text,historic_period_2015_2019_text,source_nrs_text), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom="text",x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 50, label = "Historic range (min & max)", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 29, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 33, year_number = 2020), y = 65, yend = 210, curvature = 0.20, colour = "#4477aa", alpha = 0.5, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom="text", x = compute_start_date_from_week_number(week_number = 20, year_number = 2020), y = 100, label = "Historic mean", colour = "grey20") +
  annotate(geom="curve",x = compute_start_date_from_week_number(week_number = 22, year_number = 2020), xend = compute_start_date_from_week_number(week_number = 25, year_number = 2020), y = 115, yend = 270, curvature = 0.20, colour = "grey20", linetype="dashed", arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=30, label = "2021", hjust = 0, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=30, label = "2022", hjust = 0, colour = "grey40") +
  NULL

fig_place_of_death_home_2015_2019_historic_period %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_place_of_death_home_2015-2019_range", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 18, height = 14, dpi = 300, units = "cm")
```

@@

```{r}
(fig_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)
fig_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

```{r}
fig_weekly_proportion_of_death_by_place <-
  bind_rows(
    ## current data
    merged_deaths_overall %>%
      filter(age == "all" & sex == "all"),
    ## 2015-2019
    weekly_deaths_2015_2019_overall %>%
      mutate(year = as.integer(year)) %>%  # remove the 'average' figures
      drop_na(year) %>%
      rename(deaths_all_causes = number_of_deaths)
  ) %>%
  group_by(year, week_number, place_of_death == "All") %>%
  mutate(proportion_of_deaths = deaths_all_causes / sum(deaths_all_causes)) %>%
  ungroup %>%
  filter(place_of_death != "All") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(
    data = ., aes(x = week_number, y = proportion_of_deaths, colour = place_of_death)
  ) +
  geom_line() +
  facet_wrap(~year) +
  scale_colour_viridis_d() +
  scale_x_continuous(breaks = scales::breaks_extended(6)) +
  scale_y_continuous(breaks = scales::breaks_extended(6), label = function(x) scales::percent(x = x, accuracy = 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
  ) +
  labs(
    x = "Week number", y = "Weekly proportion of deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
  ) +
  NULL

fig_weekly_proportion_of_death_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_weekly_proportion_of_death_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 20, height = 15, dpi = 300, units = "cm")
```

@@

```{r}
fig_cumulative_weekly_proportion_of_deaths_by_place <-
  cumulative_deaths_by_place_by_year %>%
  filter(place_of_death != "All") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(data = ., aes(x = week_number, y = proportion_cumulative_deaths, group = place_of_death, colour = place_of_death)) +
  geom_line() +
  facet_wrap(~year) +
  scale_colour_viridis_d() +
  scale_fill_viridis_d() +
  scale_x_continuous(breaks = scales::breaks_extended(6)) +
  scale_y_continuous(breaks = scales::breaks_extended(6), label = function(x) scales::percent(x = x, accuracy = 1)) +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
  ) +
  labs(
    x = "Week number", y = "Cumulative proportion of place of death (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
  ) +
  NULL

fig_cumulative_weekly_proportion_of_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_cumulative_weekly_proportion_of_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 20, height = 15, dpi = 300, units = "cm")
```
